Discussion Draft

### 2020-04-03 doc/Discussion.rmd
bad merge issue

[@ottoPaysuer2019].

@@

[@ma2015_cattle]. In *Peromyscus leucopus*, the wild mice had higher gwRR and stronger (gamma nu estimates) compared to lab raised mice of the same speicies [@peterson2019].

Interference is a fundamental aspect of the recombination landscape. Some basics of the constraints bounds of interference are known, yet there are still many unknowns regarding the relationship of genome wide recombination rate and interference strength.  The non-random spacing of crossovers along the 2D length of chromosomes is the first level of quantification of crossover interference (cite gamma COI papers). Logically a negative correlation is expected; increasing the number of crossovers across chromosomes would most logically be done by  more densely spacing crossovers along chromosomes and decreasing interference strength. This pattern has empirical support from the most species (Otto and Payseur) and fits well with  the fundamental relationship between the SC area or axis length, the physical upper limit for the number of crossovers.

Our results support an opposite pattern, a positive correlation between interference strength and genome wide recombination rates; we find support that interference strength has evolved in the two groups of male strains. There is a small number of positive correlations between genome wide recombination rate and interference strength in the literature. The within sex comparison of two breeds of cattle with different genome wide recombintion rates (Ma et al) [@ma2015_cattle], between lab and wild mice of Peromyscus leucopus, [@peterson2019].

@@

Models from Goldstein et al (review in Otto and Payseur) suggest that if this pattern is widespread interference evolves whenever increased recombination rates evolve. Perhaps a distinguishing feature of models which come to this finding is that the number of crossovers is kept constant. The space across multiple loci (veller?) or between multiple crossovers increases in a positive manner with genome wide recombination rates. Given that the empirical range of crossovers per chromosome is quite small (1-3  (Otto Payseur 2019, Stapley et al 207) and  the obligate crossover rule, the assumption of constraining the number of crossovers per chromosome fits well with empirical data.

>>>>

@@

**we find support that interference strength has evolved in the two groups of male strains. There is a small number of positive correlations between genome wide recombination rate and interference strength in the literature.**

Interference is a fundamental aspect of the recombination landscape. -- genome-wide recombination rate and chromaint compaction

Some basics of the constraints bounds of interference are known, yet there are still many unknowns regarding the relationship of genome wide recombination rate and interference strength.

Our results support an opposite pattern predicted from the 'logical model', a positive correlation between interference strength and genome wide recombination rates; The within sex comparison of two breeds of cattle with different genome wide recombintion rates  [@ma2015_cattle], between lab and wild mice of Peromyscus leucopus, [@peterson2019].

Logically a negative correlation is expected; increasing the number of crossovers across chromosomes would most logically be done by  more densely spacing crossovers along chromosomes and decreasing interference strength. This pattern has empirical support from the most species [@ottoPaysuer2019] and fits well with  the fundamental relationship between the SC area or axis length, the physical upper limit for the number of crossovers.

@@

Models from Goldstein et al (review in [@ottoPaysuer2019] suggest that if this pattern is widespread interference evolves whenever increased recombination rates evolve. Perhaps a distinguishing feature of models which come to this finding is that the number of crossovers is kept constant. The space across multiple loci (veller?) or between multiple crossovers increases in a positive manner with genome wide recombination rates. Given that the empirical range of crossovers per chromosome is quite small (1-3 [@ottoPaysuer2019, @stapley_variation_2017] and  the obligate crossover rule, the assumption of constraining the number of crossovers per chromosome fits well with empirical data.
